### 2025-10-09 ms/supp.tex
methods

\documentclass[11pt]{article}
\usepackage{amsmath,amssymb}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage[parfill]{parskip}
\usepackage[margin=1in]{geometry}
\usepackage{pdflscape}
\title{Supplementary information for ``Constrained circulation ensemble"}
\author{D. Kennedy}
\renewcommand{\thefigure}{S\arabic{figure}}
\begin{document}
\maketitle

\begin{figure}[h]
\centering
\includegraphics[width=20pc]{../figs/tbd.png}
\caption{A plot showing monsoon period temperature and precip in the CCE, LENS2, and observations, including maps}
\label{supp:drought}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics[width=20pc]{../figs/tbd.png}
\caption{A plot to validate the study design, specifically zonal winds}
\label{supp:winds}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics[width=40pc]{../figs/supp/scatter_ET_P_allpoints.pdf}
\caption{JJA precipitation vs. evapotranspiration, as in main text Figure 2, but instead of only showing the drought years, here we present all the data from the CESM2 large ensemble.}
\label{supp:scatter}
\end{figure}

\end{document}