### 2025-11-05 ms/supp.tex

\documentclass[11pt]{article}
\usepackage{amsmath,amssymb}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage[parfill]{parskip}
\usepackage[margin=1in]{geometry}
\usepackage{pdflscape}
\title{Supplementary information for ``Constrained circulation ensemble"}
\author{D. Kennedy}
\renewcommand{\thefigure}{S\arabic{figure}}
\begin{document}
\maketitle

\begin{figure}[h]
\centering
\includegraphics[width=30pc]{../figs/supp/gpcp_prec_map.png}
\caption{2020 GPCP JAS precipitation anomalies relative to the 1981-2020 climatology.}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics[width=30pc]{../figs/supp/cce_anom_map.png}
\caption{CCE JAS precipitation anomalies relative to the CESM2-LE climatology.}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics[width=20pc]{../figs/supp/prec_hists.png}
\caption{JAS precipitation statistics are quite similar between the CESM2-LE and the GPCP reanalysis product. TK add the LENS2 mean, and should add `2020' to CCE}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics[width=40pc]{../figs/supp/p_and_et_quantiles.png}
\caption{P and ET response to climate change varies by soil moisture quantile. E.g. dry quantiles see reduced ET in the future, and wet quantiles see enhanced ET. }
\label{supp:quantiles}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics[width=20pc]{../figs/hists/VP_hist.png}
\caption{Show that 2090 CCE is super dry, plotting histograms of MFC and 2m VP.}
\label{supp:dryness}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics[width=40pc]{../figs/supp/scatter_ET_P_allpoints.pdf}
\caption{JJA precipitation vs. evapotranspiration, as in main text Figure 2, but instead of only showing the drought years, here we present all the data from the CESM2 large ensemble.}
\label{supp:scatter}
\end{figure}

\end{document}